---
title: "Class11"
name: "Tori Ondo"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents
Download CSV file - Move this CSV file into your RStudio project
```{r}
pdbstats <- read.csv("Data Export Summary.csv", row.names=1)
```
lets look at the table
```{r}
library(knitr)
kable(pdbstats)
```

determine the percentage of structures solved by X-Ray and Electron Microscopy. From the website what proportion of structures are protein?
```{r}
nstru <- sum(pdbstats$Total)
nstru
percent <- round(pdbstats$Total / nstru*100, 2) #proportion of structures that are each type (XRAY...)
```

A1. There are `r percent[1]` % X-ray structures and `r percent[3]` % EM structures in the PDB database as of `r Sys.Date()`

```{r}
#this adds the colom of percent into the table
nstats <- pdbstats
nstats$percent <- percent
kable(nstats)
```


The proportion of protein entries
```{r}
npro <- sum(pdbstats$Proteins)
propro <- round(npro/nstru*100, 2) #proportion of sturucture that are proteins 
```

A1B. The proportion of protein entries is `r propro` %


```{r}
library(bio3d)
# to see list of functions avaialable use    lbio3d()

pdb <- read.pdb("1hsg")
pdb
```

```{r}
attributes(pdb)
# Print a subset of $atom data for the first two atoms  
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]

# Note that individual $atom records can also be accessed like this 
pdb$atom$elety[1:2]

# Which allows us to do the following 
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab="B-factor")
```

Q7) What type of R object is pdb$atom?
```{r}
str(pdb$atom)
```

```{r}
# Print a summary of the coordinate data in $xyz
pdb$xyz

# Examine the row and column dimensions  
dim(pdb$xyz)

# Print coordinates for the first two atom
pdb$xyz[ 1, atom2xyz(1:2) ]
```

```{r}
# Select all C-alpha atoms (return their indices)  
ca.inds <- atom.select(pdb, "calpha") 
ca.inds
```

```{r}
# Print details of the first few selected atoms  
head( pdb$atom[ca.inds$atom, ] )

# And selected xyz coordinates  
head( pdb$xyz[, ca.inds$xyz] )
```

```{r}
# Select chain A 
a.inds <- atom.select(pdb, chain="A") 
 
# Select C-alphas of chain A 
ca.inds <- atom.select(pdb, "calpha", chain="A") 
 
# We can combine multiple selection criteria to return their intersection 
cab.inds <- atom.select(pdb, elety=c("CA","CB"), chain="A", resno=10:20)
```


## Aligning multiple stuructures
```{r}
# Download some example PDB files 
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")  
files <- get.pdb(ids, split = TRUE)
```

```{r}
# Extract and align the chains we are interested in  
pdbs <- pdbaln(files, fit = TRUE)
 
# Print to screen a summary of the 'pdbs' object  
pdbs
```

```{r}
# find invariant core
core <- core.find(pdbs)
# superimpose all structures to core
pdbs$xyz = pdbfit(pdbs, core)
# Perform PCA
pc.xray <- pca(pdbs)
```

look at the results
```{r}
plot(pc.xray)
```

```{r}
 # Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```

