---
title: "class08"
author: "Tori Ondo"
output: github_document
date: "10/25/2018"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means Clustering

Our first example with **kmeans()** function

```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

k <- kmeans(x, centers = 2, nstart =20)
k
```
 Q. How many points are in each cluster?
      there are 30 points in each cluster
 Q. What ‘component’ of your result object details
     - cluster size?: k$size
     - cluster assignment/membership?: k$cluster
     - cluster center?: k$centers
```{r}
k$size
k$cluster
k$centers
```
     
Plot x colored by the kmeans cluster assignment and
     add cluster centers as blue points
```{r}
palette(c("orange", "green"))
plot(x, col=k$cluster)
points(k$centers, col="blue", pch = 20, cex = 3)
```
     
 Q. Repeat for k=3, which has the lower tot.withinss?
```{r}
k2 <- kmeans(x, centers = 2, nstart =20)
k3 <- kmeans(x, centers = 3, nstart =20)
k4 <- kmeans(x, centers = 4, nstart =20)
k5 <- kmeans(x, centers = 5, nstart =20)

k2$tot.withinss
k3$tot.withinss
k4$tot.withinss
k5$tot.withinss
plot(c(2, 3, 4,5), c(k2$tot.withinss,k3$tot.withinss, k4$tot.withinss, k5$tot.withinss))
```
 
 
 
##Hierarchical Clustering

Trying **hclust()* function. This needs a distance matrix as input!

```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)

#draw a line
abline(h=8, col="red")

#cut by height
cutree(hc, h=8)
```


### A more 'real' example

Step 1. Generate some example data for clustering

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
#         (just so we can compare to hclust results)
palette(c("orange", "blue", "magenta"))
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)

```
```{r}
# Q. Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters
dist2 <- dist(x)
hc3 <- hclust(dist2)
plot(hc3)

grp.3 <- cutree(hc3, k=3)
plot(x, col=grp.3, pch=20)

```

```{r}
# Q. How does this compare to your known 'col' groups?
#use the table() function !!

table(cutree(hc3, k=3))
table(col)


#this shows that col splits it into 3 even groups of 50 while hclust split it into groups of 59, 49, and 42
```


##PCA
```{r}
# Initialize a blank 100 row by 10 column matrix
mydata <- matrix(nrow=100, ncol=10)

# Lets label the rows gene1, gene2 etc. to gene100
rownames(mydata) <- paste("gene", 1:100, sep="")

# Lets label the first 5 columns wt1, wt2, wt3, wt4 and wt5
#   and the last 5 ko1, ko2 etc. to ko5 (for "knock-out")
colnames(mydata) <- c( paste("wt", 1:5, sep=""),
                       paste("ko", 1:5, sep="") )

# Fill in some fake read counts
for(i in 1:nrow(mydata)) {
wt.values <- rpois(5, lambda=sample(x=10:1000, size=1)) 
ko.values <- rpois(5, lambda=sample(x=10:1000, size=1))
mydata[i,] <- c(wt.values, ko.values)
}
head(mydata)
```

```{r}
## lets do PCA - need to use "t" becuase samples should be rows
pca <- prcomp(t(mydata), scale=TRUE)
 
attributes(pca)

# A basic PC1 vs PC2 2-D plot -- use name$x to plot
plot(pca$x[,1], pca$x[,2])


```

How much variance from original data is captured?
```{r}
# Variance captured per PC
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

pca.var.per

#Make a scree plot!
barplot(pca.var.per, main="Scree Plot",
        xlab="Principal Component", ylab="Percent Variation")
```

Lets make our PC plot more helpful
```{r}
# A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```

identify(pca$x[,1], pca$x[,2], labels=colnames(mydata)) 
this would let you add labels to individual points on the graph within the console

```{r}
# Lets focus on PC1 as it accounts for > 90% of variance
loading_scores <- pca$rotation[,1]
summary(loading_scores)

gene_scores <- abs(loading_scores)
# Sort by magnitudes from high to low
gene_score_ranked <- sort(gene_scores, decreasing=TRUE)
## Find the names of the top 5 genes
top_5_genes <- names(gene_score_ranked[1:5])

## Show the scores (with +/- sign)
pca$rotation[top_5_genes,1]
```



## Class Worksheet


