---
title: "class13"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


work with structure 1HSG
```{r}
library(bio3d) 
file.name <- get.pdb("1hsg")
```
Read this into R and examine composition
```{r}
hiv <- read.pdb(file.name)
hiv
```
 Q1: What is the name of the two non protein resid values in this structure? 
 ~HOH and MK1
 What does resid correspond to and how would you get a listing of all reside values in this structure?

## Split into protein and ligand files

we will use the 'trim.pdb()' function to split our structure

```{r}
prot <- trim.pdb(hiv, "protein")
lig  <- trim.pdb(hiv, "ligand")

#now write these into their own pdb file
write.pdb(prot, file="1hsg_protein.pdb")
write.pdb(lig, file="1hsg_ligand.pdb")
```
 
##Docking with Vina
we ran this in terminal:
~/Downloads/autodock_vina_1_1_2_mac/bin/vina --config config.txt --log log.txt

##Read docking results
we got a file all.pdbqt that we need to make into a PDB format
```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)

write.pdb(res, "results.pdb")
```

```{r}
ori <- read.pdb("1hsg_ligand.pdbqt")
rmsd(ori, res)
```


##Section 3
Normal Mode Analysis (NMA)
Normal mode analysis (NMA) is one of the major simulation techniques used to probe large- scale motions in biomolecules. Typical application is for the prediction of functional motions in proteins.
Normal mode analysis (NMA) of a single protein structure can be carried out by providing a PDB object to the function nma(). In the code below we first load the Bio3D package and then download an example structure of hen egg white lysozyme (PDB id 1hel) with the function read.pdb(). Finally the function nma() is used perform the normal mode calculation:

```{r}
pdb <- read.pdb("1HEL")
modes <- nma(pdb) 
plot(modes, sse=pdb)

# Visualize NMA
mktrj(modes, mode=7, file="nma_7.pdb")
```

